---
title: 'Identify species found in Palau EEZ'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ohibc/src/templates/ohibc_hdr1.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = TRUE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, # fig.path = 'figs/',
                      echo = TRUE, message = FALSE, warning = FALSE)

library(raster)
library(data.table)
library(sf)

source('https://raw.githubusercontent.com/oharac/src/master/R/common.R')
  ### includes library(tidyverse); library(stringr); 
  ### dir_M points to ohi directory on Mazu; dir_O points to home dir on Mazu

source('common_fxns.R')
```

# Summary

Having identified cells in the Palau EEZ, using the 100 km^2^ cell and EEZ map as a quick check, identify species found in Palau EEZ.  Copy those spp rasters over to Github for more universal access.

## Identify Palau EEZ cells

Palau EEZ for OHI is region 8.  This can be updated later using the Palau EEZ shapefile but should be a good start for now.

```{r use Palau EEZ polygons to identify regions}

palau_eez_sf <- read_sf('spatial/Shapefiles_PICRC_DFZ/Palau_EEZ.shp') %>%
  st_transform(gp_proj4)

palau_nms_sf <- read_sf('spatial/Shapefiles_PICRC_DFZ/PNMS.shp') %>%
  st_transform(gp_proj4)

palau_terr_sf <- read_sf('spatial/Shapefiles_PICRC_DFZ/Territorial_Seas_12M.shp') %>%
  st_transform(gp_proj4) %>%
  st_zm()

cell_id_rast <- raster('spatial/cell_id_rast.tif')

palau_eez_rast <- cell_id_rast %>%
  mask(palau_eez_sf) %>%
  trim()
palau_nms_rast <- cell_id_rast %>%
  mask(palau_nms_sf) %>%
  crop(palau_eez_rast)
palau_terr_rast <- cell_id_rast %>%
  mask(palau_terr_sf) %>%
  crop(palau_eez_rast)

palau_cells_df <- data.frame(
  eez = values(palau_eez_rast),
  nms = values(palau_nms_rast),
  terr = values(palau_terr_rast)) %>%
  gather(zone, cell_id, na.rm = TRUE) 

write_csv(palau_cells_df, 'spatial/palau_cell_ids.csv')

```

## Get all the species in each cell using the original raster extracts

Pull the species IDs from the rasterized species maps.  For each map (as a `.csv` file), check whether the `cell_id`s in the Palau EEZ match `cell_id`s within the species map.  If so, flag it as local, and copy to a folder in Github.

``` {r get_spp_in_palau_eez}

palau_spp_copied_flag <- 'spatial/palau_spp_copied.csv'

if(!file.exists(palau_spp_copied_flag)) {
  
  palau_eez_cells <- read_csv('spatial/palau_cell_ids.csv') %>%
    filter(zone == 'eez') %>%
    .$cell_id
  
  spp_rasts <- list.files(file.path(dir_o_anx, 'spp_rasters'), full.names = TRUE)

  copy_if_palau <- function(rast, eez_cells) {
    rast_cells <- read_csv(rast) %>%
      filter(cell_id %in% eez_cells)
    if(nrow(rast_cells) == 0) {
      message('No Palau overlap for ', basename(rast))
    } else {
      write_csv(rast_cells, file.path('spp_rasts', basename(rast)))
    }
  }
  
  parallel::mclapply(spp_rasts, FUN = copy_if_palau, 
                     eez_cells = palau_eez_cells, 
                     mc.cores = 20)

  write_csv(data.frame(copied_flag = TRUE), palau_spp_copied_flag)
}

```

## Identify species in NMS and territorial waters

By design, all species copied over are found in the EEZ.  Here we identify the subsets that fall within the NMS and within the territorial waters, respectively.

```{r}

spp_rasts <- list.files('spp_rasts', pattern = '.csv$', full.names = TRUE)
spp_cells <- parallel::mclapply(spp_rasts, FUN = read_csv, mc.cores = 20) %>%
  setNames(basename(spp_rasts)) %>%
  bind_rows(.id = 'rast') %>%
  mutate(iucn_sid = str_extract(rast, '[0-9]+') %>% as.integer) %>%
  select(-rast)

palau_cells_df <- read_csv('spatial/palau_cell_ids.csv')
palau_nms_df <- palau_cells_df %>%
  filter(zone == 'nms')
palau_terr_df <- palau_cells_df %>%
  filter(zone == 'terr')

spp_eez <- spp_cells %>%
  group_by(iucn_sid) %>%
  summarize(area = 100 * n(),
            zone = 'eez_km2')

spp_nms <- spp_cells %>%
  filter(cell_id %in% palau_nms_df$cell_id) %>%
  group_by(iucn_sid) %>%
  summarize(area = 100 * n(),
            zone = 'nms_km2')

spp_terr = spp_cells %>%
  filter(cell_id %in% palau_terr_df$cell_id) %>%
  group_by(iucn_sid) %>%
  summarize(area = 100 * n(),
            zone = 'terr_km2')

spp_palau <- bind_rows(spp_eez, spp_nms, spp_terr) %>%
  spread(zone, area, fill = 0)

write_csv(spp_palau, 'data/palau_spp_maps.csv')

```

``` {r attach spp_taxa_info}

spp_palau <- read_csv('data/palau_spp_maps.csv')

taxa_info_file <- 'data/palau_spp_taxa_info_from_api_2019-2.csv'
if(!file.exists(taxa_info_file)) {
  spp_ids <- spp_palau$iucn_sid %>% unique()
  spp_taxa_url <- 'http://apiv3.iucnredlist.org/api/v3/species/id/%s?token=%s'
  spp_taxa_df <- mc_get_from_api(spp_info_url, spp_ids, api_key, delay = 1) %>%
    clean_df_names()
  write_csv(spp_taxa_df, taxa_info_file)
}

spp_taxa_df <- read_csv(taxa_info_file)

spp_habs <- read_csv('data/spp_marine_from_api_2019-1.csv') %>%
  filter(iucn_sid %in% spp_palau$iucn_sid) %>%
  mutate(hab = str_split(habs, ', ')) %>%
  unnest(hab) %>%
  left_join(read_csv('raw/iucn_hab_categories.csv'), by = c('hab' = 'code'))

spp_info_all <- spp_palau %>%
  left_join(spp_taxa_df, by = 'iucn_sid') %>%
  left_join(spp_habs, by = 'iucn_sid') %>%
  transmute(iucn_sid,
            eez_km2,
            nms_km2,
            terr_km2,
            sciname = scientific_name,
            kingdom = str_to_sentence(kingdom),
            phylum = str_to_sentence(phylum),
            class = str_to_sentence(class),
            order = str_to_sentence(order),
            family = str_to_sentence(family),
            genus,
            comname = main_common_name,
            category,
            criteria,
            population_trend,
            depth_upper,
            depth_lower,
            depth_class = max_depth,
            hab,
            habitat)

write_csv(spp_info_all, 'output/palau_spp_info.csv')

```

``` {r}

spp_risk <- read_csv('data/iucn_risk_current_2019-2.csv') %>%
  arrange(cat_score) %>%
  mutate(cat = fct_inorder(cat))

palau_spp <- read_csv('output/palau_spp_info.csv') %>%
  select(iucn_sid, eez_km2, nms_km2, terr_km2) %>%
  distinct() %>%
  gather(zone, area, ends_with('km2')) %>%
  left_join(spp_risk, by = c('iucn_sid')) %>%
  filter(!is.na(cat)) %>%
  filter(area > 0)

ggplot(palau_spp, aes(x = cat, fill = cat)) +
  geom_bar() +
  theme_classic() +
  scale_fill_manual(values = c('green4', 'yellow3', 'orange2', 'red3', 'red4', 'grey60')) +
  labs(x = 'IUCN conservation status',
       y = 'Log (count)',
       fill = 'IUCN risk') +
  scale_y_log10() +
  facet_wrap( ~ zone)

```


``` {r old stuff, eval = FALSE}

spp_ranges   <- read_csv(file.path(dir_data, sprintf('iucn_spp_range_area_%s.csv', api_version)),
                         col_types = 'dd__') %>%
  distinct()

### make a dataframe of species risk and regional risk
spp_risk_all <- spp_risk %>%
  mutate(iucn_rgn = 'global') %>%
  bind_rows(spp_risk_rgn) %>%
  left_join(spp_ranges, by = c('iucn_sid')) %>%
  select(iucn_sid, sciname, iucn_rgn, cat_score, range_km2) %>%
  mutate(range_km2 = round(range_km2))

cells_info <- read_csv(cell_summary_file, col_types = 'd__cd____id') %>%
  distinct()

lme_to_rgn <- read_csv(file.path(dir_git, 'spatial/iucn_rgn_to_lme.csv')) %>%
  rename(rgn_name = iucn_rgn) %>%
  distinct()

taxa_cells_info <- read_csv(taxa_cells_file, col_types = 'ddcc') %>%
  left_join(spp_risk_all, by = 'iucn_sid') %>%
  mutate(spp_gp = tolower(spp_gp)) %>%
  left_join(cells_info, by = c('cell_id')) %>%
    ### add info from rasters, including lme_id
  left_join(lme_to_rgn, by = c('lme_id')) %>%
    ### add region names to filter out regional assessments
  mutate(rgn_name = ifelse(is.na(lme_id), 'global', rgn_name),
         priority = ifelse(rgn_name == 'global', 100, priority)) %>%
    ### LME cells already tagged 'global'; fix non-LME cells  and set low-ranking priority
  filter(iucn_rgn == rgn_name) %>%
  group_by(cell_id, iucn_sid) %>%
  filter(priority == min(priority)) %>%
    ### for each spp in each cell, choose the obs with highest-ranking priority
  ungroup()

spp_cells_info <- taxa_cells_info %>%
  select(cell_id, iucn_sid, sciname, spp_gp, iucn_rgn, 
         cat_score, range_km2, cell_group, depth, spp_max_depth = max_depth)

write_csv(spp_cells_info, file.path(dir_git, 'data_explore', 'spp_cells_scores.csv'))

```



