---
title: 'Quick biodiversity risk maps of Palau EEZ'
author: "*Compiled on `r date()` by `r Sys.info()['user']`*"
output: 
  html_document:
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: yes
    number_sections: true
    theme: cerulean
    highlight: haddock
    includes: 
      in_header: '~/github/ohibc/src/templates/ohibc_hdr1.html'
  pdf_document:
    toc: true
---

``` {r setup, echo = TRUE, message = FALSE, warning = FALSE}

knitr::opts_chunk$set(fig.width = 6, fig.height = 4, # fig.path = 'figs/',
                      echo = TRUE, message = FALSE, warning = FALSE)

library(raster)
library(data.table)
library(sf)

source('https://raw.githubusercontent.com/oharac/src/master/R/common.R')
  ### includes library(tidyverse); library(stringr); 
  ### dir_M points to ohi directory on Mazu; dir_O points to home dir on Mazu

source('common_fxns.R')
```

# Summary

Using Palau EEZ shapefile (and NMS etc) create quick maps showing biodiversity risk, species richness, etc.  These will be based on the output rasters from the biodiversity risk project rather than created anew for this project.

```{r load Palau shapefile}
### read in shapefiles (in WGS84) and transform to GP?
# p_eez_sf <- read_sf('spatial/Shapefiles_PICRC_DFZ/Palau_EEZ.shp') %>%
#   st_transform(gp_proj4)
# p_nms_sf <- read_sf('spatial/Shapefiles_PICRC_DFZ/PNMS.shp') %>%
#   st_transform(gp_proj4)
# p_land_sf <- read_sf('spatial/Shapefiles_PICRC_DFZ/MainIsland_Baselines.shp') %>%
#   st_transform(gp_proj4)

### let's just read in and keep in WGS84.
p_eez_sf  <- read_sf('spatial/Shapefiles_PICRC_DFZ/Palau_EEZ.shp')
p_nms_sf  <- read_sf('spatial/Shapefiles_PICRC_DFZ/PNMS.shp')
p_land_sf <- read_sf('spatial/Shapefiles_PICRC_DFZ/MainIsland_Baselines.shp')
### get extent of the Palau EEZ for cropping rasters
extent_p_eez <- extent(p_eez_sf)
### get the bbox
extent_sf <- st_as_sfc(st_bbox(p_eez_sf), forceMulti = TRUE)
```

```{r load rasters}

dir_spp_risk <- '../spp_risk_dists/_output'

### these are all in Gall-Peters equal area projection - reproject to WGS84...
risk_un_rast   <- raster(file.path(dir_spp_risk, 'mean_risk_raster_comp.tif')) %>%
  projectRaster(crs = crs(p_eez_sf), method = 'ngb')
risk_rr_rast   <- raster(file.path(dir_spp_risk, 'mean_rr_risk_raster_comp.tif')) %>%
  projectRaster(crs = crs(p_eez_sf), method = 'ngb')
threat_un_rast <- raster(file.path(dir_spp_risk, 'pct_threat_raster_comp.tif')) %>%
  projectRaster(crs = crs(p_eez_sf), method = 'ngb')
threat_rr_rast <- raster(file.path(dir_spp_risk, 'sr_rr_pct_threat_raster_comp.tif')) %>%
  projectRaster(crs = crs(p_eez_sf), method = 'ngb')
n_spp_rast     <- raster(file.path(dir_spp_risk, 'n_spp_risk_raster_comp.tif')) %>%
  projectRaster(crs = crs(p_eez_sf), method = 'ngb')
n_rr_rast      <- raster(file.path(dir_spp_risk, 'sr_rr_risk_raster_comp.tif')) %>%
  projectRaster(crs = crs(p_eez_sf), method = 'ngb')

### get land polygon
land_poly <- read_sf('spatial/ne_10m_land/ne_10m_land_no_casp.shp') %>%
  st_crop(risk_un_rast)

```

```{r}
rast_to_df <- function(rast, val = 'value') {
  rast_df <- rasterToPoints(rast) %>%
    as.data.frame() %>%
    setNames(c('long', 'lat', val))
}

plot_palau_sf <- function(rast, scale_fill) {
  rast_plot <- ggplot() +
    ggtheme_map() +
    geom_raster(data = rast_to_df(rast), aes(long, lat, fill = value)) +
    scale_fill + ### a function object
    # scale_fill_risk() +
    geom_sf(data = p_eez_sf, color = 'cyan', fill = NA) +
    geom_sf(data = p_nms_sf, color = 'red', fill = NA) +
    geom_sf(data = p_land_sf, alpha = .5, color = 'grey40', fill = 'grey60') +
    coord_sf(datum = NA)
  return(rast_plot)
}

plot_global_sf <- function(rast, scale_fill) {
  rast_plot <- ggplot() +
    ggtheme_map() +
    geom_raster(data = rast_to_df(rast), aes(long, lat, fill = value)) +
    scale_fill +
    geom_sf(data = land_poly, aes(geometry = geometry), 
            fill = 'grey96', color = 'grey40', size = .10) +
    geom_sf(data = extent_sf, alpha = .5, color = 'red', fill = NA) +
    # geom_sf(data = p_land_sf, alpha = .5, color = 'grey40', fill = 'grey60') +
    coord_sf(datum = NA)
  return(rast_plot)
}


```

### Unweighted biodiversity risk

```{r}
# ggplot() +
#   ggtheme_map() +
#   geom_raster(data = rast_to_df(risk_un_rast), aes(long, lat, fill = value)) +
#   scale_fill_risk() +
#   geom_sf(data = p_eez_sf, color = 'cyan', fill = NA) +
#   geom_sf(data = p_nms_sf, color = 'red', fill = NA) +
#   geom_sf(data = p_land_sf, alpha = .5, color = 'grey40', fill = 'grey60') +
#   coord_sf(datum = NA) +
#   theme(legend.title = element_blank())

pplot <- plot_palau_sf(risk_un_rast %>% crop(extent_p_eez),
              scale_fill = scale_fill_risk())
gplot <- plot_global_sf(risk_un_rast,
              scale_fill = scale_fill_risk()) 

print(pplot)
print(gplot)
```

### Range-rarity-weighted biodiversity risk

```{r}
pplot <- plot_palau_sf(risk_rr_rast %>% crop(extent_p_eez),
              scale_fill = scale_fill_risk())
gplot <- plot_global_sf(risk_rr_rast,
              scale_fill = scale_fill_risk()) 

print(pplot)
print(gplot)

```

### Species richness

```{r}
pplot <- plot_palau_sf(n_spp_rast %>% crop(extent_p_eez),
              scale_fill = scale_fill_viridis_c())
gplot <- plot_global_sf(n_spp_rast,
              scale_fill = scale_fill_viridis_c()) 

print(pplot)
print(gplot)

# ggplot() +
#   ggtheme_map() +
#   geom_raster(data = rast_to_df(n_spp_rast), aes(long, lat, fill = value)) +
#   scale_fill_viridis_c() +
#   geom_sf(data = p_eez_sf, color = 'cyan', fill = NA) +
#   geom_sf(data = p_nms_sf, color = 'red', fill = NA) +
#   geom_sf(data = p_land_sf, alpha = .5, color = 'grey40', fill = 'grey60') +
#   coord_sf(datum = NA)
```

### Range-rarity-weighted species richness

```{r}
pplot <- plot_palau_sf(n_rr_rast %>% crop(extent_p_eez),
              scale_fill = scale_fill_viridis_c())
gplot <- plot_global_sf(n_rr_rast,
              scale_fill = scale_fill_viridis_c()) 

print(pplot)
print(gplot)

# ggplot() +
#   ggtheme_map() +
#   geom_raster(data = rast_to_df(n_rr_rast), aes(long, lat, fill = value)) +
#   scale_fill_viridis_c() +
#   geom_sf(data = p_eez_sf, color = 'cyan', fill = NA) +
#   geom_sf(data = p_nms_sf, color = 'red', fill = NA) +
#   geom_sf(data = p_land_sf, alpha = .5, color = 'grey40', fill = 'grey60') +
#   coord_sf(datum = NA)
```

### Proportion of threatened species

```{r}
pplot <- plot_palau_sf(threat_un_rast %>% crop(extent_p_eez),
              scale_fill = scale_fill_threat())
gplot <- plot_global_sf(threat_un_rast,
              scale_fill = scale_fill_threat()) 

print(pplot)
print(gplot)

# ggplot() +
#   ggtheme_map() +
#   geom_raster(data = rast_to_df(threat_un_rast), aes(long, lat, fill = value)) +
#   scale_fill_threat() +
#   geom_sf(data = p_eez_sf, color = 'cyan', fill = NA) +
#   geom_sf(data = p_nms_sf, color = 'red', fill = NA) +
#   geom_sf(data = p_land_sf, alpha = .5, color = 'grey40', fill = 'grey60') +
#   coord_sf(datum = NA)
```

### Range-rarity-weighted proportion of threatened species

```{r}
pplot <- plot_palau_sf(threat_rr_rast %>% crop(extent_p_eez),
              scale_fill = scale_fill_threat())
gplot <- plot_global_sf(threat_rr_rast,
              scale_fill = scale_fill_threat()) 

print(pplot)
print(gplot)

# ggplot() +
#   ggtheme_map() +
#   geom_raster(data = rast_to_df(threat_rr_rast), aes(long, lat, fill = value)) +
#   scale_fill_threat() +
#   geom_sf(data = p_eez_sf, color = 'cyan', fill = NA) +
#   geom_sf(data = p_nms_sf, color = 'red', fill = NA) +
#   geom_sf(data = p_land_sf, alpha = .5, color = 'grey40', fill = 'grey60') +
#   coord_sf(datum = NA)
```


